---
title: "Database creation in the sibPlot project"
author: "Marius Bottin"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
   pdf_document:
     toc: true
     toc_depth: 5
     number_sections: true
     latex_engine: xelatex

fontsize: 11pt
geometry: "left=3cm,right=3cm,top=3cm,bottom=3cm"
linkcolor: gray
urlcolor: blue
citecolor: cyan
header-includes:
  - \usepackage{colortbl}
  - \usepackage{xcolor}       
  - \usepackage{lscape}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
--- 

```{r, setup}
require(knitr)&require(RPostgreSQL)&require(formatR)&require(kableExtra)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
opts_chunk$set(cache=F,fig.path="../../Fig/phytosoc_read",tidy='styler',cache.rebuild = F,formatSQL = TRUE, size='scriptsize')
options(knitr.kable.NA = '---')
```


# DataBase Management Systems

DataBase Management System are pieces of software which may be used in order to build the "database" part of an information system.
It is important not to confuse with the DBMS based on SQL (Structured Query Language), with SQL itself which is the informatic language used to interact with some of the DBMS.

In the *sibPlot* project, and the packages that we will build, we will integrate 2 SGDB based on SQL:

1. postgreSQL with the postgis extension
2. SQLite with the spatialite extension

The rationale behind this choice is that SQLite is a lightweight, simple file-based SGDB which might be a very reasonable choice for any biologist who want to create his database in a computer, without worrying about rights and access on the different part of the database.
However SQLite is not a sufficient solution for an organization which want to build a centralized database with clear rights, access and permissions protocols, that is where postgreSQL becomes a better choice.

Both the SGDB are based on SQL, in slightly different flavors, which may be managed through R, thanks to the *RPostgreSQL* and *SQLite* packages.
Connections are particular objects in R, that can be created through the *DBI* packages.
This makes all the management through R very easy, because the connection can be made outside of the R package we will develop, and the package will just have to check whether the connection is a SQLite or PostgreSQL one in case there are differences in the languages, which should not be often, but might happen, particularly in the case of spatial command through Spatialite or postgis...


The functions we will create should not make the configuration of the database servers, even though it might be useful to have functions reading them.
Indeed, there might be some security issues and therefore the configuration should be made by people knowing what they are doing.
The people who knows what they are doing in SGBD usually know SQL, and Bash, and should not need a R-package to simplify the process.

So what will do the package is helping organize the data and simplifying the workflow between local files and the database

# Creating the database

In SQLite, the database is all contained in a file, the process is quite simple.
On the other hand, in postgreSQL, creating the database is a much more complex process, with the need of a superuser to create the database...

So, in order to begin here, users will need in postgreSQL to define the roles and to have access to a superuser (or at least to one user having the createDB permission), and another user afterward for the writing/reading operations (in order not to make everything with a superuser, which is not a good idea).

## Process in pure SQL


## Interfacing the database creation with R



# Using dm for representing a database



```{r}
require(dm)
eval(conFlo_ra)
(dm_object<-dm_from_src(db,schema=c("main","taxon","vecmaps")))
dm_object%>%
  dm_draw
```

# Schemas

## Schema issues

Usually I would have separated the tables in many schemas, but there are 2 main issues (none of them is unsolvable though):

1. SQLite does not accept schemas (but can have 2 databases/files working at the same time if needed)
1. Many software (including the *dm* package) do not work well with tables located in different schemas, but we might make them work by writing some patches in this 

So my point here is that we should keep the separations between tables at a minimum for now...

All tables in a same schema will force us to choose well the names of the tables, but it might be for the best and could make it easier for people who are starting their journey toward SQL.
I believe we should only separate the data which has no interaction between each other, but I am still doubting even that:

* All the real data in one place
* All the specs in another place 
* all the authorization/processes/installation/administrations in another place (what might be tricky afterward is to link the person from the projects to the persons having authorizations)

# Tables





