---
title: "Input specification creation and management for the data from the Dry Forest plots from Colombia (red BST-COL)."
author: "Marius Bottin"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
   pdf_document:
     toc: true
     toc_depth: 5
     number_sections: true
     latex_engine: xelatex
always_allow_html: true
fontsize: 11pt
geometry: "left=3cm,right=3cm,top=3cm,bottom=3cm"
linkcolor: gray
urlcolor: blue
citecolor: cyan
header-includes:
  - \usepackage{colortbl}
  - \usepackage{xcolor}       
  - \usepackage{lscape}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
--- 


***************

```{r, setup, message=F, warning=F, include=F}
require(knitr)&require(RPostgreSQL)&require(formatR)&require(kableExtra)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
opts_chunk$set(cache=F,fig.path="../../Fig/input_spec_BST",tidy='styler',cache.rebuild = F,formatSQL = TRUE, size='scriptsize',
               echo=T, message=T, warning=T)
options(knitr.kable.NA = '---')
source("../functions/extract_function.R")
source("../functions/source_rmd.R")
```


**Note** there are structural deficiencies in this format. Indeed it is very dificult to find some real functional foreign keys since there is a difference between census0 and recruitment.

I have developped the functions for managing the formats taking this format into account.
Check file [specs.R](../functions/specs.R).

```{r}
require(RPostgreSQL)
sib_adm <- dbConnect(PostgreSQL(), dbname = "sib_plot", user = "sib_adm")
```

# Reinitializing Spec schema

```{r, code=readLines("../scripts/mariusTmp/reinitializeSpecSchema.r")}
source("../scripts/mariusTmp/reinitializeSpecSchema.r",max.deparse.length = Inf)
```

# Inserting BST_csv 0.1

```{r, code=readLines("../scripts/mariusTmp/reinitializeBST_csv0.1.r")}
source("../scripts/mariusTmp/reinitializeBST_csv0.1.r",max.deparse.length = Inf)
```

# Extracting useful functions
```{r}
source("../functions/extract_function.R")
```

## Regex test

The function test_regex_multi allows to find the correspondance of a character vectors against multiple regex (and is used for recognizing fields in a tables as the ones corresponding in the formats).

```{r code= extractFunction("../functions/specs.R", "test_regex_multi")}
eval(parse(text = extractFunction("../functions/specs.R", "test_regex_multi")))
```

## Check New Format

The function checkNewFormat is a R-function which allows to verify the integrity of a new input format, without much interactions with the spec schema of the database. It returns a cleaned list of tables that may be used for the new format integration in the database.


```{r code= extractFunction("../functions/specs.R", "checkNewFormat")}
eval(parse(text = extractFunction("../functions/specs.R", "checkNewFormat")))
```

## Prepare SQL queries with identifiers

The function prepareQueryId is a function which help prepare queries in the case some of the parameters are identifiers (of table or fields). Only the PostgreSQL version of a connection works for now.
It returns an argument list that may be used with do.call and dbSendQuery.

```{r code=extractFunction("../functions/specs.R","prepareQueryId")}
eval(parse(text = extractFunction("../functions/specs.R","prepareQueryId")))
```

## Insert New format

The function insertNewInFormat works with the result of checkNewFormat as an argument and allows to insert the new format specification in the spec schema of the database.

```{r code=extractFunction("../functions/specs.R","insertNewInFormat")}
eval(parse(text = extractFunction("../functions/specs.R","insertNewInFormat")))
```

## Preparing new format
### Get Input spec
The function getInputSpec allows to extract a format from the spec schema of the database.

```{r code=extractFunction("../functions/specs.R","getInputSpec")}
eval(parse(text = extractFunction("../functions/specs.R","getInputSpec")))
```

The function listInputFormats allows to list the input formats in a given database

```{r code=extractFunction("../functions/specs.R","listInputFormats")}
eval(parse(text = extractFunction("../functions/specs.R","listInputFormats")))
```


### Managing versions

The functions maxVersion and indentVersions allow to manage versions, in general, but here allows more precisely to indent and compare the versions of the input formats


```{r code=extractFunction("../functions/specs.R","maxVersion")}
eval(parse(text = extractFunction("../functions/specs.R","maxVersion")))
```

```{r code=extractFunction("../functions/specs.R","indentVersion")}
eval(parse(text = extractFunction("../functions/specs.R","indentVersion")))
```

### Preparing new input format

Preparing a new format using the extraction of an existing one is made through the function prepareInputFormat, which writes the prepared files in a given directory


```{r code=extractFunction("../functions/specs.R","prepareInputFormat")}
eval(parse(text = extractFunction("../functions/specs.R","prepareInputFormat")))
```




# Integrating BST_csv 0.2
```{r}
eval(parse(text=extractFunction("../functions/specs.R","checkNewFormat")))
eval(parse(text=extractFunction("../functions/specs.R","insertNewInFormat")))
```

```{r}
sib_adm <- dbConnect(PostgreSQL(), dbname = "sib_plot", user = "sib_adm")
spec <- checkNewFormat(conn_adm = sib_adm, mainFile = "../inputSpec/Bst_csv/BST_csv_0.2_format.csv", typeFile = "csv")
insertNewInFormat(spec,sib_adm)
```

# Integrating BST_csv 0.3

```{r}
spec <- checkNewFormat(conn_adm = sib_adm, mainFile = "../inputSpec/Bst_csv/BST_csv_0.3_format.csv", typeFile = "csv")
insertNewInFormat(spec,sib_adm)
```


# Preparing files for future version of BST_csv

```{r}
#specNew <- prepareInputFormat(sib_adm, "BST_csv", author="marius", "minor", dirName="../inputSpec/Bst_csv/", typeFile="csv")
```



```{r} 
dbDisconnect(sib_adm)
```

