---
title: "Management of general Excel functions for reading, analysing and writing excel files"
author: "Marius Bottin"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
   pdf_document:
     toc: true
     toc_depth: 5
     number_sections: true
     latex_engine: xelatex
always_allow_html: true
fontsize: 11pt
geometry: "left=3cm,right=3cm,top=3cm,bottom=3cm"
linkcolor: gray
urlcolor: blue
citecolor: cyan
header-includes:
  - \usepackage{colortbl}
  - \usepackage{xcolor}       
  - \usepackage{lscape}
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
--- 


***************

```{r, setup, message=F, warning=F, include=F}
require(knitr)&require(RPostgreSQL)&require(formatR)&require(kableExtra)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
opts_chunk$set(cache=F,fig.path="../../Fig/XL",tidy='styler',cache.rebuild = F,formatSQL = TRUE, size='scriptsize',
               echo=T, message=T, warning=T)
options(knitr.kable.NA = '---')
source("../functions/extract_function.R")
source("../functions/source_rmd.R")
```

# Separating consecutive tables in an excel worksheet

The idea is to read different tables from an excel worksheet, separated from blank lines, it is used in the spec reading from Roy's documents.


```{r, code=extractFunction("../functions/readingFile.R",nameFunction = "xl_sep_table")}
eval(parse(text=extractFunction("../functions/readingFile.R",nameFunction = "xl_sep_table")))
```


# Repeating merged cells in an excel file

In the package *openxlsx*, there is an option that is supposed to do just what we try to do here, but it seems not to work.

The idea is that when cells are merged in the excel format, we should be able to read them as a table where the values are repeated in all merged files.

At the beginning I went through *openxlsx* internal methods to extract all the functions which would allow to do that, but I realized that the workbooks read in R have style objects which may allow to extract directly the merged cells, without the need of going through complex internal functions. 
So I modified the way to do it, passing directly by the workbook object.


Let's take the example of a file called "vegetacion paramo perija (1).xlsx" (a file including various information about phytosociological relevés from Perija from O. Rangel)

```{r}
wb <- openxlsx::loadWorkbook("../../docFito/vegetacion paramo perija (1).xlsx")
```

The first function is a function which allows to extract the merged cells from a worksheet, given the workbook object and the name of the sheet

```{r, code=extractFunction("../functions/XL.R","getCellMerge")}
eval(parse(text=extractFunction("../functions/XL.R","getCellMerge")))
```

Here is the result we can obtain from the workbook that we read:

```{r}
sheets <- names(wb)
(mergedCells <- lapply(sheets, function(w,s){getCellMerge(w,s)},w = wb))
```


As you can see, that gives us the differents ranges of cells which are merged in the excel sheets.
Since those are given in a Excel-adresses type, we need a function to transform them as numeric addresses of rows and columns:

```{r, code=extractFunction("../functions/XL.R","xlAd2rowCol")}
eval(parse(text=extractFunction("../functions/XL.R","xlAd2rowCol")))
```

```{r}
xlAd2rowCol("AZ52")
```

Now we make a function for excel address ranges:

```{r, code=extractFunction("../functions/XL.R","treatCellRange")}
eval(parse(text=extractFunction("../functions/XL.R","treatCellRange")))
```


With the example of merged cells obtained before, here is the result of this function:

```{r}
lapply(mergedCells,treatCellRange)
```


The last step is to make a large general function which replaces all merged cells by the value, repeated for all cells!
The function takes as arguments:

* *fileXl*: the file (and its path)
* *getWorkbook*: a boolean telling whether we want the function to return a modified workbook (otherwise it returns a list with the data of the worksheet, with repeated values for the merged cells)
* *writeFile*: whether we want the function to write a file with repeated values in all unmerged cells
* *dos_unMerged*: path of the folder where to write the file
* *overwrite*: whether to overwrite the file which would be already present in the folder (with a name corresponding with the name of the original file with "_unmerged")


```{r, code=extractFunction("../functions/XL.R","xl_fillMerged")}
eval(parse(text=extractFunction("../functions/XL.R","xl_fillMerged")))
```

